---
title: 'Lab #12'
author: "Nick Gembs"
date: "12/14/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Create a functions that determines whether a testValue is inside an interval, or outside an interval. It should take two arguments: One is the testValue. The second argument should be a vector of length two that represents an interval. It should return TRUE, if the testValue is outside the interval.Otherwise it returns FALSE.

```{r}

outside = function( testValue, interval){
    return(testValue < min(interval) | testValue > max(interval))
}

```



```{r}
############# 
rMysteryGenerator <- function(n, shift = 0) {
    a <- qchisq(runif(n), df = 2)
    b <- pchisq(a, df = 2)
    c <- pbeta(rbeta(n, shape1 = 1, shape2 = 2), shape1 = 1, shape2 = 2)
    d <- sqrt(-2 * log(b)) * cos(2 * pi * c) + punif(2) + shift
    e <- pnorm(d)
    return(d)
}
############# 
```



2. rMysteryGenerator is a random number generator. Decide if the distribution meets the standards for conducting a two-sided test about a mean.

```{r}

samp = rMysteryGenerator(10000, shift = 0)
hist(samp)
```
Yes


3. Hypothesize a value for the mean of the distribution.

The mean appears to be around 1.

4. You will use a sample size of 10 to conduct a two sided hypothesis test with a  significance level of 0.05. Determine the critical values for such a test.

```{r}

alpha = .05
n = 10
cv = qt(alpha/2, df=n-1, lower.tail=F)
cv
```


5. Take a sample of size 10 from this distribution. Decide whether your hypothesis for the mean appears to be correct or not.

```{r}

samp =  rMysteryGenerator(10, shift = 0)
tResults = t.test(samp, mu=1, alternative="two.sided")
tResults
str(tResults)
testStat = tResults$statistic
outside(testStat, c(-cv, cv))
```



6. Repeat the test 10000 times. Determine the fraction of the tests where you would reject the null hypothesis, and what fraction do you fail to reject the null hypothesis.

```{r}

reps = 10000
holder = rep(NA,reps)

for (i in 1:reps){
    samp =  rMysteryGenerator(n, shift = 0)
    tResults = t.test(samp , mu=1 , alternative = "two.sided")
    testStat = tResults$statistic
    holder[i] = outside(testStat, c(-cv,cv))
}
mean(holder)
```



7. If a test is perfect, what should the number that was produced be?

0


8. The TA will inform you of the actual mean. What does your result form part 6 mean?


The actual mean is 1. Our null hypothesis is true, our test has a .0498 type 1 error rate.


9. The **shift** argument shifts the mean of the distribution. Thus, if the original mean is 12 and shift is equal to .5, the distribution used to generate the sample  has a mean of 12.5. Repeat 6-9 with a shift of 0.1.

```{r}

reps = 10000
holder = rep(NA,reps)

for (i in 1:reps){
    samp =  rMysteryGenerator(n, shift = .1)
    tResults = t.test(samp , mu=1 , alternative = "two.sided")
    testStat = tResults$statistic
    holder[i] = outside(testStat, c(-cv,cv))
}
mean(holder)

```


10. Increase the size of the shift, does the probability of a Type II error increase or decrease?

```{r}
reps = 10000
holder = rep(NA,reps)

for (i in 1:reps){
    samp =  rMysteryGenerator(n, shift = .15)
    tResults = t.test(samp , mu=1 , alternative = "two.sided")
    testStat = tResults$statistic
    holder[i] = outside(testStat, c(-cv,cv))
}
mean(holder)

```
The probability of a Type II error decreases.

11. Determine how big **shift** must be so that the probability of a Type II error is approximately 0.10.

```{r}
possibleshifts = seq(0,5,by=.1)
ln = length(possibleshifts)
holdMeanHolder = rep(NA, ln)

for ( i in 1:ln){
    reps = 1000
    holder = rep(NA,reps)

    for (j in 1:reps){
        samp =  rMysteryGenerator(n, shift = possibleshifts[i])
        tResults = t.test(samp , mu=1 , alternative = "two.sided")
        testStat = tResults$statistic
        holder[j] = outside(testStat, c(-cv,cv))
    }
    holdMeanHolder[i] = mean(holder)
}    

plot(possibleshifts, holdMeanHolder, type="l")

pos = which.max(holdMeanHolder >= .9)
possibleshifts[pos]
```
With a sample size of 10, the shift away from the mean has to be at least 1.2 to reduce the type II error rate to .1






