---
title: "Lab #6"
subtitle: "Section 404"
author: "Nick Gembs"
date: "10/21/2020"
output:
  word_document: default

---


## Examples

This code is to be used in the examples. This is the code that was created in the examples from Lab #5.




```{r , echo=TRUE}
#Addem is a function that will add together the elements of a vector
#Addem takes a single argument X with a default value of 0
Addem<-function( X = 0 ){
  total<-0
  for(i in X){
    total<-total+i
  }
  return(total)
}


# SimulateDieRollMatrix is a function that simulates rolling a fair die with at least 2 sides.
# SimulateDieRollMatrix will produce a matrix with ROLLS * REPETITIONS values.
# Each column will hold the results of rolling a simulated die ROLLS number of times.
# There will be REPETITIONS columns. For example, you want to roll a die 10 times, and record the resuls. 
# This will constitute a single column. Suppose you want to repeat this 3 times. Then you will have three columns, each of ten rolls.
# SimulateDieRollMatrix takes three arguments. SIDES is the number of sides on the die. ROLLS is the number of rolls. 
# REPETITIONS is the number of times you repear making ROLLS rolls.
SimulateDieRollMatrix<-function( REPETITIONS , ROLLS , SIDES = 2 ){
  die<-1:SIDES
  totalRolls <- ROLLS*REPETITIONS
  dieMat <- matrix( NA ,nrow=ROLLS,ncol=REPETITIONS)
  for( i in 1:REPETITIONS){
    outcomes <- sample(die , size=ROLLS , replace=TRUE)
    dieMat[,i] <- outcomes
    }
  return(dieMat)
}

# ADDCOLS is a function that takes each column of a matrix, sums the elements, and returns a vector with these sums.
# The function takes 3 arguments
# MATRIX is the matrix whose rows need to be added together
# NROW is the number of rows in MATRIX
# NCOL is the number of columns in MATRIX
ADDCOLS<-function(MATRIX,NROW,NCOL){
  ADDED<-rep(NA,times=NCOL)
  for(i in 1:NCOL){
    ADDED[i]<-Addem(MATRIX[,i])
  }
  return(ADDED)
}


```
 

 
1. Create a function called **HISTROLLS**. **HISTROLLS** creates a histogram. The histogram will approximate the distribution of the sum of rolling a **SIDES** sided fair die **ROLLS** number of times. **SIDES** and **ROLLS** are two arguments. **REPETITIONS** will be the third argument. It will determines the number of sums that will be used in creating the histogram. 

```{r fig.height=5, fig.width=5, out.width=200, echo=TRUE}
HISTROLLS <- function(REPETITIONS, ROLLS, SIDES = 2) {
  simulatedData <- SimulateDieRollMatrix(REPETITIONS, ROLLS, SIDES)
  sums <- ADDCOLS(simulatedData, ROLLS, REPETITIONS)
  
  hist(sums, main = paste ("Number of Rolls= ", ROLLS))
}

HISTROLLS(REPETITIONS = 3000, ROLLS = 50, SIDES = 6)
```
 

 
2. Create a function called **GRAPHS**. **GRAPHS** will produce a series of histograms. Each histogram will represent the distribution of summing the results of rolling a **SIDES** sided die **ROLLS** number of times. The histograms will appear side-by-side. It will take 4 arguments. **SIDES** which details the number of sides on the simulated die. **ROLLS** is a vector containing the number of times the simulated die is rolled.
 
```{r fig.height=4, fig.width=12, out.width=400, echo=TRUE}
GRAPHS = function(REPETITIONS = 2000, ROLLS = 1, SIDES = 6) {
  numberofgraphs = length(ROLLS)
  par(mfrow = c(1,numberofgraphs))
  for (i in 1:numberofgraphs) {
    HISTROLLS(REPETITIONS, ROLLS[i], SIDES)
  }
  par(mfrow = c(1,1))
}
GRAPHS(REPETITIONS = 5000, ROLLS = 1:5)
```

 
3. CHANGE **eval=FALSE** in the next code chunk to **eval=TRUE** once you have completed the previous examples.
 
```{r fig.width=12,fig.height=4,out.width=400, echo=TRUE, eval=TRUE}
GRAPHS( REPETITIONS=3000 ,  ROLLS=c(2,4,6) , SIDES=8    )
```


\newpage

## Problems

Each Problem is a modified version of the examples given in lab.  The following code is MY version of the answers for Lab #5. It is not YOUR version of the answers for lab#5.  You may use my code for Lab #6, or you may use your version. If you want to use your code from Lab #5, delete my code. If you want to use mine for Lab #6, that is fine also. Your answers for Lab #5 should be your work, and not mine. 


```{r , echo=TRUE}


AverageM<-function( X = 0 ){
  n<-length(X)
  average<-Addem(X)/n
  return(average)
}

SIMBUCKETS<-function(REPETITIONS, BALLSELECTED, DIST="uniform"){
  set.seed(1)
  ballMat<-matrix(NA, nrow = BALLSELECTED , ncol = REPETITIONS)
    for(i in 1:REPETITIONS){  
      if(DIST=="uniform"){
        outcomes<-runif(BALLSELECTED)
      }else {
        outcomes<-rchisq( BALLSELECTED , df = 4.3 )
      }
      ballMat[,i]<-outcomes
    }
  return(ballMat)
}


EX<-SIMBUCKETS(3,2,DIST="uniform")
EX

AverageCOLS<-function(MATRIX,NROW,NCOL){
  AVE<-rep(NA,times=NCOL)
  for(i in 1:NCOL){
    AVE[i]<-AverageM(MATRIX[,i])
  }
  return(AVE)
}

RESULTS<-AverageCOLS( EX , NROW = 2, NCOL = 3)
RESULTS

```

\newpage

4. Make a function called **HISTDIST**. It should take three arguments:**NROWS**, **NCOLS**, **DIST**.  **HISTDIST** should create a histogram of the averages of the columns of a  **NROWS** by **NCOLS** matrix from a given distribution. Use the previous distribution options for **DIST**:  "chisq2" or "uniform". Enter two extra arguments into the hist() command beyond just the vector of averages. The first extra argument is called **xlim**. It will depend on **DIST** It should be set to c(0,10)  for **DIST**= "chisq2",  and c(0,1) for **DIST** = "uniform". (Use an if else statement)The second extra argument should be breaks=40. This will be the same regardless of the option for **DIST**. 

```{r fig.height=3,fig.width=4,out.height=200, echo=TRUE}

HISTDIST <- function(NROWS, NCOLS, DIST = "uniform") {
  mat <- SIMBUCKETS(NCOLS, NROWS, DIST) 
  Average <- AverageCOLS(mat,NROWS, NCOLS)
  if (DIST == "chisq2") {
    xlim <- c(0,10)
  } else {
    xlim <- c(0,1)
  }
  hist(Average, xlim = xlim, breaks = 40 )
}
HISTDIST(NROWS = 5, NCOLS = 5000, DIST = "uniform")

```
 


5. Make a function **GRAPHDISTS**. It should take three arguments: **REPETITIONS**, **BALLSSELLECTED**, **DIST**. **REPETITIONS** should have a default value of 2000. **BALLSSELLECTED** should default to c(10,20,30). **DIST** should default to "uniform".  **GRAPHDISTS** should produce a series of histograms. Each histogram should simulate the distribution of the averates of selected **BALLSSELLECTED** numbered balls from a bucket of numbered balls labeled **DIST**. Each histogram should be created using a different element from **BALLSSELLECTED**. Reset the plotting region to 1x1 at the end of this.


```{r fig.height=3,fig.width=12,out.height=200, echo=TRUE}
GRAPHDISTS <- function(REPETITIONS = 2000, BALLSSELECTED = c(10,20,30), DIST = "uniform") {
  numberofgraphs = length(BALLSSELECTED)
  par(mfrow = c(1,numberofgraphs))
  for (i in 1:numberofgraphs) {
    HISTOGRAM <- HISTDIST(BALLSSELECTED[i], REPETITIONS, DIST)
  }
  par(mfrow = c(1,1))
}
GRAPHDISTS(REPETITIONS = 5000)

          
                    
```




6. At the end, you should be able to run these three lines of code, and produce 3 plots with 6 histograms each. REMOVE the command eval=FALSE when you are ready to knit.

```{r fig.height=3,fig.width=12,out.height=200, echo=TRUE}
GRAPHDISTS(REPETITIONS = 2000 , BALLSSELECTED = c(1,2,5,10) , DIST = "uniform" )
GRAPHDISTS(REPETITIONS = 2000 , BALLSSELECTED = c(1,2,5,10) , DIST = "chisq2" )

```

 



