---
title: ''
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Stat Computing - Exercises 02 - Gradebook

There is a dataset in `datasets/grade_book.csv` containing simulated
grades for Statistical Computing. From the exercises directory, you
can read in the dataset by running

```{r}
dat <- read.csv("../grade_book.csv")
str(dat)
```
Complete the following exercises related to the grade book.

1. Randomly generate with replacement a birth date for each student between
   2001-01-01 and 2005-12-31. Print out how many unique birth dates there
   are (year-month-date) and how many unique birthdays there are (month-date).
   Is this surprising?
```{r}
dim(dat)

bday = (as.Date(runif(50, min = 0, max = 1825), origin = as.Date("2001-01-01")))
bdayshort = format(bday, format = "%m/%d")
length(unique(bday))
b = length(unique(bdayshort))
b
```

There are 50 unique (year-month-date) birthdays and `r b` unique (month-date) 
birthdays. This makes sense because there are 1825 unique birthday options with 
year while only 365 without year.


2. Add the birthdate column to the data frame in 3 different ways: using
   `cbind`, `$`, and `[[]]`.
```{r}
way1 = cbind(dat,birthdate = bday)

#way2
dat$birthdate = bday

#way3
dat[["birthdate"]] = bday

```


3. Print out the column names.
```{r}
names(dat)
```


4. Remove the birthdate column and show that it's gone.
```{r}
dat$birthdate = NULL
names(dat)
```


5. Print out which column number has the netIDs.
```{r}
which( names(dat)=="netID" )
```


6. Print out the column numbers that contain lab grades. You might find the
   `grep` function useful.
```{r}
grep('lab', names(dat))
```


7. Print out the column numbers that contain exercise grades.
```{r}
grep('ex', names(dat))
```


8. What happens when you try to convert the data frame to a matrix with
   `as.matrix`?
```{r}
matdat <- as.matrix(dat)
```

It automatically converts all of the data types to string because a matrix has 
to have uniform data type.

9. Extract the exercise columns and convert to a matrix. Why does this work as intended?
```{r}
datmat = as.matrix(dat[grep('ex', names(dat))])
```

This works as intended because the rule of uniform data type of a matrix is not 
broken, so R does not auto convert.

10. Add a column to the data frame containing each student's average exercise
   grade.  Treat missing values as a grade of 0. You can do this in a couple of
   lines with `rowSums` or `rowMeans`. Exercises are out of 20. Print out the average
   exercise grades for the first 10 students.
```{r}

datmat <- replace(datmat, is.na(datmat), 0)

5*rowMeans(datmat)[1:10]
dat$exAVG = 5*rowMeans(datmat, na.rm =TRUE)
```


11. Calculate each student's exercise average again, this time using the
    average of the non-missing values. Print out the average exercise grades
    for the first 10 students.
```{r}
dat1 <- read.csv("../grade_book.csv")
datmat = as.matrix(dat[grep('ex', names(dat1))])
5*rowMeans(datmat, na.rm =TRUE)[1:10]


```

12. Print out the number of missing exercises for each exercise.
```{r}
dat1 <- read.csv("../grade_book.csv")
datmat = as.matrix(dat1[grep('ex', names(dat1))])
colSums(is.na(datmat))

```


13. Calculate each student's lab average, and add to the data frame. Labs are out of 10.
    Print out the average lab grades for the first 10 students.
```{r}

datmat = as.matrix(dat[grep('lab', names(dat))])

10*rowMeans(datmat, na.rm =TRUE)[1:10]

#datmat <- replace(datmat, is.na(datmat), 0)
dat$labAVG = 10*rowMeans(datmat, na.rm =TRUE)


```


14. Using the formula in the syllabus, add a column containing each student's
    overall numeric grade. Treat missing assignments as 0. Project is out of 100.
```{r}




dat <- replace(dat, is.na(dat), 0)

dat$numericgrade = dat$labAVG*.2 + dat$exAVG*.6 + dat$project*.2
head(dat)

```


15. Using the guidelines in the syllabus, add a column containing each
    student's letter grade.
```{r}
lam = c()
for(i in 1:50){
   if(dat[i,"numericgrade"] >= 93){
      lam[i] = "A"
   } else if (dat[i,"numericgrade"] >= 90){
      lam[i] = "A-"
   } else if (dat[i,"numericgrade"] >= 87){
      lam[i] = "B+"
   } else if (dat[i,"numericgrade"] >= 83){
      lam[i] = "B"
   } else if (dat[i,"numericgrade"] >= 80){
      lam[i] = "B-"
   } else if (dat[i,"numericgrade"] >= 77){
      lam[i] = "C+"
   } else if (dat[i,"numericgrade"] >= 73){
      lam[i] = "C"
   } else if (dat[i,"numericgrade"] >= 70){
      lam[i] = "C-"
   } else if (dat[i,"numericgrade"] >= 67){
      lam[i] = "D+"
   } else if (dat[i,"numericgrade"] >= 63){
      lam[i] = "D"
   } else if (dat[i,"numericgrade"] >= 60){
      lam[i] = "D-"
   } else {
      lam[i] = "F"
   }
}

dat$lettergrade = lam

head(dat)
```


16. Print out the netID, numeric average, and letter grade for the top 10
    scorers. You may want to look at the `order` function.

```{r}
tops = dat[order(dat$numericgrade, decreasing=TRUE),c("netID","numericgrade","lettergrade")]
tops[1:10,]
```

