---
title: ""
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Statistical Computing - Exercises 04 - Ithaca Weather

The US government, via its agencies NOAA and NCEI, maintains the US
Climate Reference Network, a collection of 139 high quality weather
stations whose purpose is to monitor atmospheric and soil
conditions. One of them is located right here outside of Ithaca, along
Route 38.

We will be formatting and analyzing the hourly Ithaca data from 2021
and 2022, with the goal of building a model for predicting afternoon
temperatures. 

1. Read the dataset `datasets/CRNH0203-2021-NY_Ithaca_13_E.txt` into R as a
   data frame and add column names according to the
   [**documentation**](https://www.ncei.noaa.gov/pub/data/uscrn/products/hourly02/).
   Make sure you inspect your data frame and do any necessary cleaning.

```{r}
# Putting read.table & strsplit on separate lines for readability creates a
#knitting error. All other lines formatted.

dat <- read.table("C:/Users/Nick/Documents/GitHub/statcomp2023/datasets/CRNH0203-2021-NY_Ithaca_13_E.txt")
columns = strsplit("WBANNO UTC_DATE UTC_TIME LST_DATE LST_TIME CRX_VN LONGITUDE LATITUDE T_CALC T_HR_AVG T_MAX T_MIN P_CALC SOLARAD SOLARAD_FLAG SOLARAD_MAX SOLARAD_MAX_FLAG SOLARAD_MIN SOLARAD_MIN_FLAG SUR_TEMP_TYPE SUR_TEMP SUR_TEMP_FLAG SUR_TEMP_MAX SUR_TEMP_MAX_FLAG SUR_TEMP_MIN SUR_TEMP_MIN_FLAG RH_HR_AVG RH_HR_AVG_FLAG SOIL_MOISTURE_5 SOIL_MOISTURE_10 SOIL_MOISTURE_20 SOIL_MOISTURE_50 SOIL_MOISTURE_100 SOIL_TEMP_5 SOIL_TEMP_10 SOIL_TEMP_20 SOIL_TEMP_50 SOIL_TEMP_100", split = " ")

colnames(dat) <- columns[[1]]



for (i in columns){
   
   holder = dat[i]
   holder = replace(holder,(dat[i] == -999.0), NA )
   holder = replace(holder,(dat[i] == -9999.0), NA )
   holder = replace(holder,(dat[i] == -99999.0), NA )

   dat[i] = holder
}


```




2. Save the data frame to your computer using both `save` and `saveRDS`. What
   are the two file sizes, and how do the sizes compare to the size of the
   original .txt file? (You can check this in the terminal with the command
   `ls -lah`)

```{r}
saveRDS(dat, file = "C:/Users/Nick/Documents/weatherdat.rds")

save(dat, file = "C:/Users/Nick/Documents/weatherdat.RData")
```
The saved files are 204 KB, which is much smaller than the 2096 KB text
document.

3. Create a data frame containing only the longitude and latitude. Save the
   result as both a .csv and a .RData file. How do the sizes compare? Is this
   surprising to you? Can you explain the results?
   
```{r}
dat1 = dat[, c("LONGITUDE", "LATITUDE")]

save(dat1, file = "C:/Users/Nick/Documents/weatherdatlonglat.RData")
save(dat1, file = "C:/Users/Nick/Documents/weatherdatlonglat2.csv")
write.csv(dat1, file = "C:/Users/Nick/Documents/weatherdatlonglat.csv")
```
Both save files are only 1 KB, this is surprising because the Rdata file should 
be smaller due to compression. This is likely happening because the stored data
is still being compressed, and then sent to csv when using the save function. 
When you directly write to a csv file, the size is much larger (179 kb). This
is because write.csv does not undergo compression.

The Rdata file is only 1 KB while the csv file is 179 KB. This is because 
RData files undergo compression while csv files only separate the values with
a comma.

4. Create a new data frame that has one row for each day. Its columns should be
   the date, 3pm air temperature (use local standard time for all times and
   T_CALC for all air temperatures), 6am temperature, 7am temperature, 8am
   temperature, and 9am temperature.
   
```{r}
dat2 <- dat[(dat["LST_TIME"] == 600) | (dat["LST_TIME"] == 700) |
               (dat["LST_TIME"] == 800) 
            | (dat["LST_TIME"] == 900) | (dat["LST_TIME"] == 1500) ,
            c("T_CALC","LST_TIME","LST_DATE")]

date <- unique(dat2["LST_DATE"])
sixAM <- dat2[dat2["LST_TIME"]==600,"T_CALC"]
sevenAM <- dat2[dat2["LST_TIME"]==700,"T_CALC"]
eightAM <- dat2[dat2["LST_TIME"]==800,"T_CALC"]
nineAM <- dat2[dat2["LST_TIME"]==900,"T_CALC"]
threePM <- dat2[dat2["LST_TIME"]==1500,"T_CALC"]

newdat<- data.frame(date,sixAM,sevenAM,eightAM,nineAM,threePM,
                    row.names = 1:365)
head(newdat)

```



5. Do a multiple regression of 3pm temperature on 6am, 7am, 8am, and 9am
   temperature and print out the summary. How accurate does the model say your
   predictions should be?

```{r}
m1 <- lm(newdat$threePM ~ newdat$sixAM + newdat$sevenAM + newdat$eightAM + 
            newdat$nineAM)
summary(m1)
```
With an F-statistic of 1081 and a p-value less than 2.2e-16, we can reject a
null hypothesis of there being no relation between 6-9 am temps and 3 pm
temp. Since the residual standard error is 3.027, the model claims that 
predictions will be within 3.027 degrees of observed ~68% (1 sd) of the time,
and within 6.054 degrees of observed ~95% (2 sd) of the time.


6. Use your model to predict 2022 3pm temperatures from morning temperatures.
   You'll have to read in the 2022 data and organize it the same way you
   organized the 2021 data. Report the root mean squared prediction error and
   make a plot of the model's prediction errors against day of year.
```{r}
dat_ <- read.table("C:/Users/Nick/Documents/GitHub/statcomp2023/datasets/CRNH0203-2022-NY_Ithaca_13_E.txt")
columns = strsplit("WBANNO UTC_DATE UTC_TIME LST_DATE LST_TIME CRX_VN LONGITUDE LATITUDE T_CALC T_HR_AVG T_MAX T_MIN P_CALC SOLARAD SOLARAD_FLAG SOLARAD_MAX SOLARAD_MAX_FLAG SOLARAD_MIN SOLARAD_MIN_FLAG SUR_TEMP_TYPE SUR_TEMP SUR_TEMP_FLAG SUR_TEMP_MAX SUR_TEMP_MAX_FLAG SUR_TEMP_MIN SUR_TEMP_MIN_FLAG RH_HR_AVG RH_HR_AVG_FLAG SOIL_MOISTURE_5 SOIL_MOISTURE_10 SOIL_MOISTURE_20 SOIL_MOISTURE_50 SOIL_MOISTURE_100 SOIL_TEMP_5 SOIL_TEMP_10 SOIL_TEMP_20 SOIL_TEMP_50 SOIL_TEMP_100", split = " ")

colnames(dat_) <- columns[[1]]

holder = dat$T_CALC
holder = replace(holder,(dat$T_CALC == -9999.0), NA )

dat_$T_CALC = holder

dat2_ <- dat_[(dat_["LST_TIME"] == 600) | (dat_["LST_TIME"] == 700) | 
                 (dat_["LST_TIME"] == 800) 
            | (dat_["LST_TIME"] == 900) | (dat_["LST_TIME"] == 1500) ,
            c("T_CALC","LST_TIME","LST_DATE")]
#dat2

date <- unique(dat2_["LST_DATE"])
sixAM <- dat2_[dat2_["LST_TIME"]==600,"T_CALC"]
sevenAM <- dat2_[dat2_["LST_TIME"]==700,"T_CALC"]
eightAM <- dat2_[dat2_["LST_TIME"]==800,"T_CALC"]
nineAM <- dat2_[dat2_["LST_TIME"]==900,"T_CALC"]
threePM <- dat2_[dat2_["LST_TIME"]==1500,"T_CALC"]

newdat_<- data.frame(date,sixAM,sevenAM,eightAM,nineAM,threePM,
                     row.names = 1:365)

predictions <- predict(m1, newdata = newdat_)

perrors = predictions-threePM
predictionerrors = abs(predictions-threePM)
rmse = sqrt(sum((predictionerrors[(is.na(predictionerrors)==F)])^2)/365)
rmse

plot(1:365, perrors, xlab = "day", ylab = "Model Prediction Error")

```
The model rmse is `r rmse`

7. Build your own model for 3pm temperatures based on the 2021 data. You can
   use any information in the data, and any kind of model you like, but you
   must use only information that's available at the 9am hour or earlier (e.g.
   you cannot use today's 2pm temperature to predict today's 3pm temperature).
```{r}
dat2 <- dat[ (dat["LST_TIME"] == 500) |(dat["LST_TIME"] == 700) | 
                (dat["LST_TIME"] == 800) 
            | (dat["LST_TIME"] == 900) | (dat["LST_TIME"] == 1500) , ]

date <- unique(dat2["LST_DATE"])
minsunrise = dat2[dat2["LST_TIME"]==500,"T_MIN"]
max9am = dat2[dat2["LST_TIME"]==900,"T_MAX"]
fourhourtempchange = max9am-minsunrise
sevenAM <- dat2[dat2["LST_TIME"]==700,"T_CALC"]
eightAM <- dat2[dat2["LST_TIME"]==800,"T_CALC"]
nineAM <- dat2[dat2["LST_TIME"]==900,"T_CALC"]
threePM <- dat2[dat2["LST_TIME"]==1500,"T_CALC"]

sevenAMrain <- dat2[dat2["LST_TIME"]==700,"P_CALC"]
eightAMrain <- dat2[dat2["LST_TIME"]==800,"P_CALC"]
nineAMrain <- dat2[dat2["LST_TIME"]==900,"P_CALC"]

nineAMSOLARAD <- dat2[dat2["LST_TIME"]==900,"SOLARAD"]
nineAMSoil <- dat2[dat2["LST_TIME"]==900,"SOIL_TEMP_5"]

rain <- sevenAMrain + eightAMrain + nineAMrain

newdat<- data.frame(date,sevenAM,eightAM,nineAM,fourhourtempchange, rain , 
                    nineAMSOLARAD, nineAMSoil ,threePM, row.names = 1:365)


m1 <- lm(newdat$threePM ~ newdat$fourhourtempchange + newdat$sevenAM + 
            newdat$eightAM + newdat$nineAM + newdat$sevenAM*newdat$nineAM
         + newdat$rain + newdat$nineAMSOLARAD + newdat$nineAMSoil)
summary(m1)
```
   

8. Test out your model on the 2022 data. Report the root mean squared
   prediction error and make a plot of the model's prediction errors against
   day of year.
   
```{r}

dat2_ <- dat_[(dat_["LST_TIME"] == 500) | (dat_["LST_TIME"] == 700) | 
                 (dat_["LST_TIME"] == 800) 
            | (dat_["LST_TIME"] == 900) | (dat_["LST_TIME"] == 1500) ,]
#dat2

date <- unique(dat2_["LST_DATE"])
minsunrise = dat2_[dat2_["LST_TIME"]==500,"T_MIN"]
max9am = dat2_[dat2_["LST_TIME"]==900,"T_MAX"]
fourhourtempchange = max9am-minsunrise
sevenAM <- dat2_[dat2_["LST_TIME"]==700,"T_CALC"]
eightAM <- dat2_[dat2_["LST_TIME"]==800,"T_CALC"]
nineAM <- dat2_[dat2_["LST_TIME"]==900,"T_CALC"]
threePM <- dat2_[dat2_["LST_TIME"]==1500,"T_CALC"]

sevenAMrain <- dat2_[dat2_["LST_TIME"]==700,"P_CALC"]
eightAMrain <- dat2_[dat2_["LST_TIME"]==800,"P_CALC"]
nineAMrain <- dat2_[dat2_["LST_TIME"]==900,"P_CALC"]

nineAMSOLARAD <- dat2_[dat2_["LST_TIME"]==900,"SOLARAD"]
nineAMSoil <- dat2_[dat2_["LST_TIME"]==900,"SOIL_TEMP_5"]

newdat_<- data.frame(date,sevenAM,eightAM,nineAM,fourhourtempchange, rain ,
                     nineAMSOLARAD, nineAMSoil ,threePM, row.names = 1:365)

predictions <- predict(m1, newdata = newdat_)

perrors = predictions-threePM
predictionerrors = abs(predictions-threePM)
rmse = sqrt(sum((predictionerrors[(is.na(predictionerrors)==F)])^2)/365)
rmse

plot(1:365, perrors, xlab = "day", ylab = "Model Prediction Error")
```
   The model rmse is `r rmse`
   
   
   
   
   
