---
title: "Exercises 3"
author: "Nick Gembs"
date: "2/6/2023"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
---
title: "Exercises 3"
author: "Nick Gembs"
date: "2/6/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Stat Computing - Exercises 03 - Billboard Charts

On canvas, there is a file called `Hot_100.csv`, which contains data from the
Billboard 100 music charts going back to 1958. Below are some exercises on
functions and subsetting that make use of the data. For all questions, include
3 examples to demonstrate calling your function with different arguments.

```{r}
dat = read.csv("C:/Users/Nick/Downloads/Hot_100.csv")
```

1. Write a function that takes in a character string indicating a date, an
   optional format argument that has a default date format. The function should
   return the #1 song and its performer during the week that contains the date
   argument. Assume that the date in the dataset is the date of the beginning
   of the week. Your function should use the default date format if a format is
   not provided, and it should use the user-supplied format if it is supplied.
   Test your function with a few different dates and different formats.
   
```{r}
topsong <- function(date, format = "%Y-%m-%d") {
  date = as.Date(date, format = format)
  weekdaychange = 6 - as.numeric(format(date, "%u"))

  date = format(date+weekdaychange, "%Y-%m-%d")
  
  return(dat[(dat$chart_date == date) & (dat$chart_position == 1),
             c("song","performer")])
}

topsong("1990-05-05")
topsong("1990-07-17" , format = "%Y-%m-%d")
topsong("1990/07/17" , format = "%Y/%m/%d")
topsong("17-07-1990" , format = "%d-%m-%Y")
topsong("28-09-1990" , format = "%d-%m-%Y")
```



2. Modify your function to add a `top` argument, indicating that the function
   should return the top `top` songs on the specified date. Make sure that the
   returned songs are listed in order according to the weekly ranking. Test
   your function by printing the top 10 songs on the day you were born.

```{r}

topsong <- function(date, format = "%Y-%m-%d", top = 1) {
  date = as.Date(date, format = format)
  weekdaychange = 6 - as.numeric(format(date, "%u"))

  date = format(date+weekdaychange, "%Y-%m-%d")
  
  result = (dat[(dat$chart_date == date) & (dat$chart_position <= top),
                c("song","performer", "chart_position")])
  
  
  return(result[order(result$chart_position),c("song", "performer")])
}

topsong("2002-07-17", top = 10)
topsong("2002-2-11", top = 5)
topsong("2022-2-10", top = 1)
topsong("2002-1-28", top = 5)
```


3. Write a function to search for matching performer names. The function should
   take in a character string as its argument, find all unique performer names
   that contain the argument, and return them.  Include an argument to
   determine whether or not to do exact matching on uppercase vs. lowercase.
   It may be a good idea to look at the documentation `grep`, `tolower`, and
   `toupper`.
   
```{r}

SearchArtist <- function(artist, MatchCase = FALSE){
  if (MatchCase == FALSE){
    return ( dat[grep(artist, dat$performer, ignore.case = TRUE, 
                      fixed = FALSE),])
  } else if (MatchCase == TRUE) {
    return ( dat[grep(artist, dat$performer, ignore.case = FALSE,
                      fixed = TRUE),])
  }
}

head(SearchArtist("Emin"))
head(SearchArtist(("Charlie Puth"), MatchCase = TRUE))
head(SearchArtist(("Joe"), MatchCase = TRUE))
```
   

4. Write a function to take in a performer name, and return all of the songs
   belonging to that performer that appeared on the Billboard list.  Use exact
   matching for performer names.
   
```{r}
artistsongs <- function (artist) {
  return ( unique(dat[grep(artist, dat$performer, ignore.case = FALSE,
                           fixed = TRUE), "song"]))
}

artistsongs("Black Eyed Peas")
cat("\n")
artistsongs("Yung Gravy")
cat("\n")
artistsongs("Foo Fighters")
```
   

5. Song statistics: write a function that takes in a song id, and returns the
   following statistics for the song: number of weeks in the top 100, number of
   weeks in the top 40, number of weeks in the top 10, number of weeks at #1,
   lowest position (i.e. best position). The return type should be a data frame
   with columns for the song name, the performer, the first week it appeared on
   the charts, and all of the statistics. 

```{r}

Song_Statistics <- function (song_ID) {
   
   song_dat = dat[dat$song_id == song_ID,]
   top_100 = length(song_dat$song)
   top_40 = length(song_dat[song_dat$chart_position <= 40 , "song"])
   top_10 = length(song_dat[song_dat$chart_position <= 10 , "song"])
   top_1 = length(song_dat[song_dat$chart_position <= 1 , "song"])
   top = min(song_dat$chart_position)
   
   result = data.frame(song = song_dat[1,"song"], 
                       performer = song_dat[1,"performer"], 
                       first_week = song_dat[1,"chart_date"] , 
                       top_100 = top_100, top_40 = top_40 ,
                       top_10 = top_10, top_1 = top_1,
                       best = top)
   
   return ((result))
  
}

Song_Statistics("PepasFarruko")
Song_Statistics("HelloAdele")
Song_Statistics("Blinding LightsThe Weeknd")

```

6. Song rankings: write a function that ranks all of the songs according to the
   statistics in the previous question. The function should take a statistic
   name as its argument, and return a data frame whose first column is the song
   name, second column is the performer, third column is the week the song
   first appeared, and the fourth column is the statistic. The data frame
   should be ordered with the best value of the statistic at the top of the
   data frame. Print out the rankings of the top 10 songs for 3 of the
   statistics.

```{r}
# Citation: Idea for code of this problem came from office hours

Rankings = function(statistic) {
   songs_id <- unique(dat$song_id)
   
   df2 = data.frame(colnames(c("song","performer", "chart_debut", "top_100",
                               "top_40", "top_10", "top_1", "best")))
   
    for ( i in seq_along(songs_id)) {
      stats2 <- Song_Statistics(songs_id[i])
      df2 <- rbind(df2, stats2)
      
   }
  
   if (statistic == "best") {
      resultdf <- (df2[order(df2[,statistic], decreasing = FALSE),c("song",
                                      "performer", "first_week", statistic)])
   } else {
      resultdf <- (df2[order(df2[,statistic], decreasing = TRUE),c("song",
                                    "performer", "first_week", statistic)])
   }
   
   finalresult <- head(resultdf, 10)
   return ( finalresult )

}

Rankings("top_40")
Rankings("top_1")
Rankings("best")

```






