---
title: "STSCI 4520 Lab 1"
author: "Nick Gembs"
date: "1/25/2023"
output:
  word_document: default
  html_document: default
---

---
title: "Introduction to R Markdown and factor variables"
output: pdf_document
---

# Lab Exercise

Complete the exercise below in a SEPARATE R Markdown file, and turn
it in as a pdf on canvas. DO NOT include the introduction to
Rmarkdown from the top of this script.

Read in the "voice_heights" data found in the github repo in
`datasets/voice_heights.csv`. The data contain the heights and voice
categories for all people in a choir. Fit a linear model to the
heights, using voice as a factor

```{r}
dat <- read.csv("C:/Users/Nick/Documents/GitHub/statcomp2023/datasets/voice_heights.csv")
head(dat)

h <- dat$height
v <- as.factor(dat$voice)
levels(v)
m1 <- lm( h ~ v )
summary(m1)
```

Note that the regression output compares the height of each voice with
the reference level.

Re-run the regression 3 more times, each time changing the reference level.
Use the results to report the t-statistic for comparing the means of
every pair of heights.

```{r}
v <- as.factor(dat$voice)
v <- relevel(v, ref = "bass")
m1 <- lm( h ~ v )
summary(m1)
```
*The regression above reports t-statistics for mean heights of choir group factor variables against a reference of bass. The t-statistic for alto vs bass is -9.961, soprano vs bass is -11.129, and tenor vs bass is -2.267*

```{r}
v <- as.factor(dat$voice)
v <- relevel(v, ref = "soprano")
m1 <- lm( h ~ v )
summary(m1)
```
*The regression above reports t-statistics for mean heights of choir group factor variables against a reference of soprano. The t-statistic for alto vs soprano is 1.065, soprano vs bass is 11.129, and tenor vs soprano is 6.987*


```{r}
v <- as.factor(dat$voice)
v <- relevel(v, ref = "tenor")
m1 <- lm( h ~ v )
summary(m1)
```
*The regression above reports t-statistics for mean heights of choir group factor variables against a reference of tenor. The t-statistic for alto vs tenor is -6.050, tenor vs bass is 2.267, and tenor vs soprano is -6.987*

You can inspect the factor variables from an `lm` object using its `xlevels`
field.

```{r}
m1$xlevels
```


```{r}
mat <- matrix(data = c(NA, 9.961,-1.065,6.050,-9.961,NA,-11.129,-2.267,1.065,11.129,NA,6.987,-6.050,2.267,-6.987,NA), ncol = 4, nrow = 4)
v <- relevel(v, ref = "alto")
colnames(mat) <- levels(v)
rownames(mat) <- levels(v)
mat
```

