---
title: "lab3"
author: "Nick Gembs"
date: "2/6/2023"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
---
title: "lab3"
author: "Nick Gembs"
date: "2/6/2023"
output: html_document
---
---
title: ""
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r load-packages, include=FALSE}
library(dplyr)
library(knitr)
```
# Exercises

1. Use `apply` to compute column sums of the matrix in the first section.

```{r}
M <- matrix( 1:12, 4, 3 )
colSums(M)
apply(X = M, MARGIN = 2, FUN = sum)
```


2. Read in the airline data and use one of the apply functions to
figure out how many missing values there are in each column of the
airline data. Make sure the output is a named vector.

```{r}
dat = read.csv(
  "C:/Users/Nick/Documents/GitHub/statcomp2023/datasets/airline_2019-07-01.csv")
sapply(dat, function(x) sum(is.na(x)))
```


3.  Use tapply to compute a matrix holding the distances between 
every pair of airports. You’ll have to read the documentation for tapply to see
how to deal with multiple factors. Print out the rows and columns for the 10 
airports with the most flights

```{r}
mat = matrix(data = 0, nrow = length(unique(dat$Origin)),
             ncol = length(unique(dat$Origin)), 
             dimnames = list(unique(dat$Origin),unique(dat$Origin)))

mat1 = tapply(dat$Distance , list(dat$Origin, dat$Dest) , mean )

flights = dat %>% group_by(Flights, Origin) %>% mutate(TotalFlights = n())

flights = flights[!duplicated(flights[,15]),]

flights = head(flights[order(flights$TotalFlights, decreasing=TRUE),], 10)

flights[,c("Origin", "TotalFlights")]

indices = c('ATL', 'ORD','DFW','DEN','CLT','LAX','SFO','IAH','PHX','LAS')

mat1[indices,indices]

```


