---
title: "stsci4520_lab7"
author: "Nick Gembs"
date: "3/8/2023"
output:
  pdf_document: default
  html_document: default
---
```{r}
# let's look at the exp function for an example
# we'll compare it to using exponentiation
n <- 1000
x <- rnorm(n)
y <- 10*rnorm(n)
e <- exp(1)

library("bench")
# spend a little time reading the documentation

# this can be alternatively called with the "cute" syntax bench::mark( exp(x) )
result <- mark( exp(x), e^x )
# this contains a table with information about how long it takes
# exp(x) is a little bit faster than e^x
result
# run this block of code a few times to get a sense of whether the results are consistent

# let's do y and compare
result <- mark( exp(y), e^y )
result

```


```{r}

# Exercise: read the documentation for "press" and use it to
# benchmark exp(x) against e^x for a range of values of n,
# such as n = 10, n = 100, n = 1000, n = 10000




# the exponential function has a series expansion:
# exp(x) = 1 + x/1! + x^2/2! + x^3/3! + ...

# let's try to beat the exp function using the series expansion
my_exp1 <- function(x){
  z <- rep(1,length(x))
  y <- x
  for(j in 1:3){
    z <- z + y/factorial(j)
    y <- y*x
  }
  return(z)
}

# this only works for small x
x <- 0.01*rnorm(n)
mark( exp(x), e^x, my_exp1(x) )
```



```{r}
n <- 1000
# Exercise: see if you can write your own version of my_exp that beats exp
my_exp2 <- function(x){
  # fill in your code here

  z <- 1 + x + x*x*((3+x)/6)
  
  return(z)
}

x <- 0.01*rnorm(n)
mark( exp(x), e^x, my_exp1(x), my_exp2(x) )


n <- 999999
x <- 0.01*rnorm(n)

t1 <- proc.time()
ans = exp(x)
t2 <- proc.time()
print("exp(x) time:")
print(t2-t1)

t1 <- proc.time()
ans1 = my_exp1(x)
t2 <- proc.time()
print("my_exp1(x) time:")
print(t2-t1)

t1 <- proc.time()
ans2 = my_exp2(x)
t2 <- proc.time()
print("my_exp2(x) time:")
print(t2-t1)


```


```{r}
# sorting strings vs factors vs numbers
hot <- read.csv("C:/Users/Nick/Downloads/Hot_100.csv")

# pay attention to the units!
mark(
  order( hot$chart_date ),
  order( as.Date( hot$chart_date ) ),
  order( as.factor(hot$chart_date) ),
  order( as.numeric(as.factor(hot$chart_date) ) )
)
```
factors and numbers return total time in milliseconds, while string and Date
return in seconds. numbers were the quickest to sort, followed by factors,
then dates, and finally, strings.


```{r}
# how much time is spent on the actual ordering?
hot_string <- hot$chart_date
hot_date <- as.Date( hot$chart_date )
hot_fact <- as.factor( hot$chart_date )
hot_number <- as.numeric( as.factor( hot$chart_date ) )
mark( order(hot_string), order(hot_date), order(hot_fact), order(hot_number) )
```
The results were different for the actual ordering. factors, dates, 
and numbers return total time in milliseconds, while string
returns in seconds. factors were the quickest to order, followed by dates,
then numbers, and finally, strings.

